{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}

### 2015-06-04 assess.Rnw
incorporated Jim's comments on revision, need to proof

Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated \citep{Cryer08}:
\begin{equation} \label{do_sin}
DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
\end{equation}
such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.  The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).  Although the representation of $DO_{die}$ as a simple sine wave does not account for lags with the solar cycle or periods of photoinhibition that may be observed in actual time series, we consider the approach sufficient for the simulations.  Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
\begin{equation} \label{do_unc_n}
DO_{unc, n} = \varepsilon_{obs, n} + \sum_{t = 1}^{n} \varepsilon_{pro, t}
\end{equation}
where observation and process uncertainty ($\varepsilon_{obs}$, $\varepsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.  Process uncertainty was estimated as a serially correlated variable using the cumulative sum of $n$ observations plus random variation added at each time step for $t = 1, ..., n$.  The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).

A tidal series was simulated based on the principal lunar semidiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.  The amplitude was set to 1 meter and centered  at 4 meters.  The tidal time series simulated \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}). Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
\begin{equation} \label{deltdo}
\frac{\delta DO_{adv}}{\delta t} = \frac{\delta DO}{\delta x} \cdot \frac{\delta x}{\delta t}
\end{equation}
where the change in \ac{DO} from advection over time $t$ is the product of the change in \ac{DO} over the change in tidal excursion $x$ and the change in tidal excursion over time.  The simulated tidal time series was used to create $DO_{adv}$ as a simple sine wave as in \cref{do_sin}.  This time series was centered at zero to simulate increasing or decreasing \ac{DO} with tidal height change.  The final time series for observed \ac{DO} was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs,fig:do_sim}).  For actual time series, changes in \ac{DO} from advection can be described generally as a function of tidal height:
\begin{equation} \label{do_advp}
DO_{adv} = f\left(H\right)
\end{equation}
The functional relationship between tidal height and changes in \ac{DO} from physical advection may vary depending on tidal regime, e.g., standing wave versus progress tides.  For weighted regression, all that is required is a variable that is statistically correlated with tide to quantify the influence on the \ac{DO} time series.  Our simulated data assumes that tidal height is directly proportional to changes in \ac{DO} and, therefore, statistically correlated.  This assumption may be invalid for actual time series but it is appropriate for evaluating the empirical qualities of the weighted regression model, as described in the previous section.

@@

Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under a range of conditions. Specifically, the simulated data varied in the relative amount of noise in the measurement ($\varepsilon_{obs}$, $\varepsilon_{pro}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2,187 weighted regressions.

The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to characterize the biological \ac{DO} time series in \cref{do_obs}. Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.  Overall, the weighted regressions produced filtered time series that were reasonable representations of the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}, results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).  The median correlation between the filtered and biological values for all time series and window widths was \Sexpr{form_fun(perf_medians['cor'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'cor'])} (very poor) to \Sexpr{form_fun(perf_rngs[2, 'cor'])} (perfect).  Mean error was \Sexpr{form_fun(perf_medians['err'])}, with values ranging from \Sexpr{form_fun(perf_rngs[1,'err'], 0, 0, 0)} (perfect) to \Sexpr{form_fun(perf_rngs[2, 'err'])} (very poor).  Simulations with very poor performance were those those with the \ac{DO} signal composed entirely of noise from observation uncertainty. As expected, simulations with no biological or tidal influence had filtered time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero).

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process error ($e_{pro}$), whereas increasing observation error contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance was not significantly affected by increasing tidal effects (i.e., increasing magnitude of $DO_{adv}$).  Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{tab:dtd_perf2,fig:err_surf2}).  Graphical summaries of model performance averaged by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.

@@

The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.  \ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is adequate for the entire water column, including the benthos.  Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}; \citealt{Kelley14}; \citetalias{RDCT14}). The predicted time series of tidal height from the regressions were used for all models to avoid issues with missing values in the observed data. The stations were generally semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).

@@

The weighted regression method was applied to the annual data for each station to obtain a filtered \ac{DO} time series for estimating metabolism.  Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey14}.  The method is used to infer net ecosystem metabolism using the mass balance equation:
\begin{equation} \label{metrate}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey14}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  This coefficient accounts for the influence of meteorological effects, such as wind, on air-sea gas exchange at the surface.  The diffusion-corrected \ac{DO} flux estimates as hourly rates of \ac{DO} change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series. The `metabolic day' was considered the period between sunsets on two adjacent calendar days.  Respiration was hourly \ac{DO} flux during night hours and net production was hourly \ac{DO} flux during day hours.  Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average \ac{DO} flux during night hours multiplied by 24. Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average daily respiration, multiplied by total sunlight time.  Net ecosystem metabolism was gross production (positive) plus total respiration (negative).  Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth at the site.  Site depth was taken from the water quality sensor and a half-meter was also added to account for approximate placement slightly off of the bottom.

The selection of half-window widths for filtering the \ac{DO} time series was based on an evaluation of results using four performance metrics.  The `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.  First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable.  Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.  The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.  We assumed that annual means would not change after filtering because long-term metabolic averages are not likely to be biased from advection (M. W. Beck unpubl.).  However, the variance is expected to be reduced because noise from advection would be removed.  Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data. Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative gross production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}. Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant cause of these values given the tidal characteristics at each site.

@@

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a percentage of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration. Anomalous values were substantially reduced for all case studies, particularly Rookery Bay and Sapelo Island.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.

>>>>

Each component of the simulated time series was created as follows.  First, the diel component, $DO_{die}$, was estimated \citep{Cryer08}:
\begin{equation} \label{do_sin}
DO_{die} = \alpha + \beta\cos\left(2\pi ft + \Phi\right)
\end{equation}
such that the mean DO ($\alpha$) was 8, amplitude ($\beta$) was 1, $f$ was 1/48 to represent 30 minute intervals, $t$ was the time series vector and $\Phi$ was the x-axis origin set for an arbitrary sunrise at 630.  The diel signal was increasing during the day and decreasing during the night for each 24 hour period and ranged from 7 to 9 mg L$^{-1}$ (as a rough approximation from the case studies below).  Although the representation of $DO_{die}$ as a simple sine wave ignores more complex dynamics associated with the photosynthesis-irradiance relationship as applied to the ecosystem, it is sufficient to address the analytical objectives of the simulations.  Uncertainty was added to the diel \ac{DO} signal as the sum of observation and process uncertainty:
\begin{equation} \label{do_unc_n}
DO_{unc, n} = \varepsilon_{obs, n} + \sum_{t = 1}^{n} \varepsilon_{pro, t}
\end{equation}
where observation and process uncertainty ($\varepsilon_{obs}$, $\varepsilon_{pro}$) were simulated as normally distributed random variables with mean zero and standard deviation varying from zero to an upper limit, described below.  Process uncertainty was estimated as a serially correlated variable using the cumulative sum of $n$ observations plus random variation added at each time step for $t = 1, ..., n$.  The total uncertainty, $DO_{unc}$, was added to the diel \ac{DO} time series to create the biological \ac{DO} time series (\cref{do_bio,fig:do_sim}).

A tidal series was simulated based on the principal lunar semidiurnal (M2) tide with a period of 12.42 hours \citep{Foreman89}.  The amplitude was set to 1 meter and centered  at 4 meters.  The tidal time series simulated \ac{DO} changes with advection, $DO_{adv}$ (\cref{do_obs_all,fig:do_sim}). Conceptually, this vector represents the rate of change in \ac{DO} as a function of horizontal water movement from tidal advection such that:
\begin{equation} \label{deltdo}
\frac{\partial DO_{adv}}{\partial t} = \frac{\partial DO}{\partial x} \cdot \frac{\partial x}{\partial t}
\end{equation}
where the change in \ac{DO} from advection over time $t$ is the product of the change in \ac{DO} over the change in tidal excursion $x$ and the change in tidal excursion over time.  The simulated tidal time series was used to create $DO_{adv}$ as a simple sine wave as in \cref{do_sin}.  This time series was centered at zero to simulate increasing or decreasing \ac{DO} with tidal height change.  The final time series for observed DO was the sum of biological \ac{DO} and advection \ac{DO} (\cref{do_obs}and \cref{fig:do_sim}). For time series from the \ac{NERRS} monitoring sites, changes in \ac{DO} associated with advection are modeled as linear function of tidal height as indicated by the parameter $\beta_{2}$ in \cref{funform}.  We can express the tidal component of \ac{DO} variations alone as:
\begin{equation} \label{do_advp}
DO_{adv} = \beta_{2}H
\end{equation}

@@

Multiple time series were simulated by varying the conditions in \cref{do_obs_all} (\cref{fig:sim_ex}) to evaluate weighted regression under different conditions. Specifically, the simulated data varied in the relative amount of noise in the measurement ($\varepsilon_{obs}$, $\varepsilon_{pro}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Half-window widths (day, hour of day, and tide height) for the weighted regressions were evaluated for each time series: time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2,187 weighted regressions.

The filtered \ac{DO} time series were compared to the simulated data to evaluate the ability of weighted regression to quantify and remove the variability due to advection ($DO_{adv}$), leaving the biological response ($DO_{bio}$) (\cref{do_obs,fig:err_surf1}).  Comparisons were made using Pearson correlation coefficients and the \ac{RMSE}.  The ability of weighted regression to isolate $DO_{bio}$, indicated by high correlation and low \ac{RMSE}, was not reduced by the introduction of advection into the simulated time series (i.e., $DO_{obs}$) because weighted regression was able to quantify and remove this variability.  On the other hand, introduction of observation uncertainty ($\varepsilon_{obs}$) decreased the ability of the model to quantify and remove advection.  Perhaps counterintuitively, increased magnitude of advection improved the model performance because a larger advection signal can be fitted empirically despite the presence of random error.  When the advection signal was small, it became difficult or impossible to quantify unless there was neither process uncertainty nor observation error.  This suggests that if the advection signal is in fact very small, it would be preferable not to quantify it and instead neglect it, as has been done previously. Model performance was not substantially affected by variation in half window widths relative to characteristics of the \ac{DO} time series (\cref{fig:err_surf2}).

@@

The weighted regression model was applied to continuous \ac{DO} time series and water level measurements from January 1\textsuperscript{st} to December 31\textsuperscript{st} 2012 at the four sites.  \ac{NERRS} sites are typically shallow and vertically-mixed such that one water quality monitor is adequate to characterize \ac{DO} for the entire water column.  Tide predictions were obtained for each site using harmonic regression applied to the sonde depth data (\texttt{oce} package in R, \citealt{Foreman89}; \citealt{Kelley14}; \citetalias{RDCT14}). The predicted time series of tidal height from the regressions were used for all models to avoid issues with missing values in the observed data. The stations were generally semidiurnal or mixed semidiurnal and net heterotrophic on an annual basis (\cref{tab:case_att}).

@@

The weighted regression method was applied to the annual data for each station to obtain a filtered \ac{DO} time series for estimating metabolism.  Ecosystem metabolism was estimated using the open-water technique \citep{Odum56} as described in \citet{Caffrey14}.  The method is used to infer net ecosystem metabolism using the mass balance equation:
\begin{equation} \label{metrate}
\frac{\partial DO}{\partial t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\partial DO$, g O$_2$ m$^{-3}$) over time ($\partial t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey14}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO} concentration, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  This coefficient accounts for the influence of meteorological effects, such as wind, on air-sea gas exchange at the surface.  The diffusion-corrected \ac{DO} flux estimates as hourly rates of \ac{DO} change were first averaged during day and night periods for each 24 hour `metabolic day' in the time series. The `metabolic day' was considered the period between sunsets on two adjacent calendar days.  Respiration was hourly \ac{DO} flux during night hours and net production was hourly \ac{DO} flux during day hours.  Total respiration (Rt) rates were assumed constant during day and night such that daily rates were calculated as the average \ac{DO} flux during night hours multiplied by 24. Daily gross production (Pg) was the average \ac{DO} flux during day hours minus the average daily respiration, multiplied by total sunlight time.  Net ecosystem metabolism was gross production (positive) plus total respiration (negative).  Finally, volumetric rates were converted to depth-integrated (areal) estimates (g O$_2$ m$^{-2}$ d$^{-1}$, \cref{tab:case_att}) by multiplying by the mean water-column depth at the site. Site depth was estimated from the pressure sensor on the water quality sonde, with a half-meter added to account for placement of the sonde just above the bottom.

The selection of half-window widths for filtering the \ac{DO} time series was based on an evaluation of results using four performance metrics.  The `optimal' window widths for each case study were those that provided the greatest measure of performance based on all four metrics.  First, the regression results were evaluated using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model, such that window widths that provided maximum reduction in correlation relative to the observed data were desirable.  Daily metabolism estimates before and after filtering were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day and respiration rates were compared to mean rates of change during the night.  The second and third performance metrics evaluated changes in the annual mean metabolism estimates and standard deviation before and after filtering.  We assumed that annual mean values would not change after filtering because long-term metabolic averages are not likely to be biased from advection. We evaluated this assumption by computing metabolism for a 12 month simulated \ac{DO} time series with and without advection.  These simulations showed that the means converge within the annual time scale, although variance of metabolism was substantially larger for the time series with advection.  Optimum window widths in this context were considered those that maintained the mean values while reducing standard deviation relative to the observed data. Finally, results were evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative gross production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}. Optimum window widths for this metric were those that provided the maximum reduction in anomalous values.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assumed that physical processes were the dominant cause of these values given the tidal characteristics at each site.

@@

The percentage of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for all sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (as a percentage of the total estimates, Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Pg', 'Value']))} (Padilla Bay) for production and \Sexpr{form_fun(min(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Rookery Bay) to \Sexpr{form_fun(max(anom_obs[anom_obs$Metab == 'Rt', 'Value']))} (Elkhorn Slough) for respiration. Anomalous values were substantially reduced for all case studies and were virtually eliminated in two of the four cases.  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (\Sexpr{form_fun(elk_per[, 'Pg_Avg'],1,1,1)} \% change from the annual mean) and respiration (\Sexpr{form_fun(elk_per[, 'Rt_Avg'],1,1,1)} \%) for Elkhorn Slough, increased mean production (\Sexpr{form_fun(pdb_per[, 'Pg_Avg'],1,1,1)} \%) and respiration (\Sexpr{form_fun(pdb_per[, 'Rt_Avg'],1,1,1)} \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Changes in mean estimates based on filtered \ac{DO} time series suggests that the weighted regression removed variation attributed to both biological and physical processes.  The implications of these undesirable results are described below.  Decreases in the standard deviation for all metabolism estimates were observed for all cases after filtering.  However, reduction in variation may also occur if the biological signal is reduced individually or in addition to physical variation.  The extent to which this reduction is related to the former should be minimal, provided that the two are statistically distinguishable.  Situations when the phasing of the tidal and solar cycles are correlated could be instances when the two are unable to be separated by the model.  Reductions in annual mean values (particularly at Elkhorn Slough) in addition to variance reduction suggests that true metabolism may have been removed causing a downward bias in the estimates.

### 2015-06-04 comments.Rnw
word formatted for resubmission
incorporated Jim's comments on revision, need to proof

A useful approach for conceptualizing instances when weighted regression may be most appropriate can be described using an ``end member'' paradigm.  These end members are the terminal points of the water mass that is horizontally advected past the \ac{DO} sensor.  One end member is typically characterized by tidal marshes during flood tide, whereas the second end member may be characterized by oxygenated oceanic waters during ebb tide.  An important distinction is made between end members that are influenced by actual biological processes that occur at each terminus and those that are not, where the former case is more likely a common occurrence at locations with large tidal amplitudes.  In the latter case, variation in dissolved oxygen at the \ac{DO} sensor is solely related to physical advection and weighted regression should work as intended provided collinearity with the solar cycle is minimal.  At sites where biological processes affect \ac{DO} at the end members, the tidal characteristics of a site and the chosen window widths are critical determinants of the ability of weighted regression to isolate the true metabolic signal.  Biological signals from diurnal or semidiurnal tides that exhibit rapid pregression with the solar cycle can be statistically isolated with relatively smaller window widths as the average (i.e., detided) signal can be quantified in fewer tidal cycles.  Conversely, sites with mixed tides that exhibit prolonged synchronicity with the solar cycle, as at Elkhorn Slough, will require larger window widths to characterize the true metabolic signal.  Issues of collinearity at mixed sites add additional challenges as previously noted.  Caution should be used with larger window widths as the potential to oversmooth and mute actual interday variation is increased.  For diurnal and semidiurnal sites, relatively small window widths that produce noticeable differences in estimates may have meaningful implications for interpreting metabolism from filtered time series.  We conclude that weighted regression has the ability to describe the true metabolic signal regardless of tidal characteristics and biological processes that vary at the end members, although chosen window widths and collinearity with the solar cycle affect its application.

>>>>

Tidal currents in estuaries can move water across a mosaic of habitat types that likely have different biological rates.  For example, the magnitudes of production and respiration, and the net metabolism may differ across salinity gradients, between open water and shallow water habitats, and inside and outside of tidal marshes and seagrass habitats.  These complex estuarine seascapes are especially prevalent at sites like those in the \ac{NERRS} network.  The interaction between spatial differences in metabolic processes and tidal advection moving water masses among habitats contributes to tidal artifacts in open water metabolism calculations.  Advection alone, such as in an open coastal environment might actually generate little distortion of the metabolic signal within the \ac{DO} record.

The objective of the weighted regression method is the remove or reduce variations associated with tides, allowing for an improved estimate of the average metabolic rates between the ``end members'' that define the tidal excursion about the \ac{DO} sensor.  The method works because in most tidal regimes, tidal cycles pregress fully relative to the solar cycle within several weeks.  Pregression occurs quickly with semidiurnal tides and slightly more slowly with diurnal tides.  At time scales longer than this period, \ac{DO} variations associated with metabolism can be separated statistically from \ac{DO} variations associated with tides.  Locally-weighted regression is necessary because both the average metabolic rates associated with the solar cycle and the horizontal \ac{DO} gradients created by spatial differences in metabolism are variable in time.  The temporal scale and the choice of window width can be based on the time required for pregression of tides relative to the solar cycle.

Problems may arise in applying the method in certain situations, apparently most common with mixed tides, where complete phase pregression occurs across a seasonal time scales.  For example, \citet{Nidzieko14} showed that the higher high tide, which floods high marsh habitats, can occur at night in Elkhorn Slough throughout summer causing both real and apparent changes in metabolism.  Large errors in standard open water metabolism calculations result because tide is generally rising during afternoon and falling in the early hours of the day, amplifying the true metabolic signal.  Because the tide and solar cycle are correlated, however, our method ascribes all variability to the tide, resulting in equally unsatisfactory and large underestimates of rates.  Weighted regression could be used, but only at a very long time scale, where the benefits of using it are not as great.  As \citet{Nidzieko14} showed, these complex patterns are best quantified and understood with the benefit of a more detailed data set.

### 2015-06-04 discussion.Rnw
incorporated Jim's comments on revision, need to proof

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.  Further, time averages of metabolism estimates using the filtered \ac{DO} time series were significantly different than those using the observed data, particularly for relatively long periods of observation depending on location. These results suggest that previous estimates may not accurately reflect true metabolic signals if the effects of tidal variation confound biological signals in observed \ac{DO} time series.  The regression method and open-water technique may also have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within \ac{NERRS}.  Extension to stratified systems is possible, provided time series across depth strata are available.  The open-water technique assumes the water column is vertically mixed or that estimates from a surface sensor only apply to the upper layer in a stratified system \citep{Staehr10,Kemp12}.  Time series from different depths would be required to estimate metabolic rates that are vertically-integrated, followed by additional validation to ensure rates agree with expectations.  In particular, application of the open-water method to each depth stratum would require an estimate of gas exchange across the pycnocline, whereas the technique implemented herein relied on a model for air-sea gas exchange at the surface.  Meaningful metabolic estimates could be obtained using such an approach, following the application of weighted regression to remove tidal influences.

@@

For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was presumably caused by physical advection.  In particular, filtering reduced the frequency of anomalous metabolism estimates, such as negative production and positive respiration \citep{Needoba12}.  It is important to recognize, however, that `normal' estimates (positive production and negative respiration) may still include significant bias due to physical advection.  For example, a large decrease in \ac{DO} at night may appear `normal' even if inflated by a tidal effect.  In some cases, apparent tidal effects could reflect real ecosystem processes. For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime because inundation of salt marshes during the night followed by draining during a falling tide during the day lead to larger respiration values.  Resolving these patterns may simply require better data.  For the weighted regression approach, sustained periods of synchrony between the solar and tidal cycles presents a problem because collinearity between the two diminishes performance.  Moreover, \ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.  For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}.  Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible.  The open water metabolism method has generally been applied at the time scale of one day, rather than hourly.  Therefore, weighted regression may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.

>>>>

The weighted regression approach was developed to improve estimates of ecosystem metabolism by removing variation associated with tidal change in observed \ac{DO} time series.  The application to simulated \ac{DO} time series with known characteristics and extension to continuous monitoring data from selected \ac{NERRS} sites suggested the approach can isolate and remove a significant amount of variation in observed \ac{DO} from tidal change.  In application to \ac{DO} time series from \ac{NERRS} sites, the method reduces apparent interday variation in metabolic rates.  This may indicate that such interday variations result substantially from artifacts of advection, rather than real drivers of metabolic processes. This hypothesis could be investigated further by applying our proposed method while collecting measurements of metabolism and causal factors at a relatively high frequency.  The regression method and open-water technique may have broad appeal for application in estuaries that are shallow and vertically-mixed, such as those within the \ac{NERRS}. Extension to stratified systems may be possible with additional data and model development.

@@

For all case studies, weighted regression was generally successful in reducing the variation in the \ac{DO} time series that was presumably caused by physical advection.  In particular, filtering reduced the frequency of anomalous metabolism estimates, such as negative production and positive respiration \citep{Needoba12}.  It is important to recognize, however, that `normal' estimates (positive production and negative respiration) may still include significant bias due to physical advection.  For example, a large decrease in \ac{DO} at night may appear `normal' even if inflated by a tidal effect.  In some cases, apparent tidal effects could reflect real ecosystem processes. For example, \citet{Nidzieko14} observed that net metabolism at Elkhorn Slough was strongly heterotrophic during spring tides that occurred at nighttime because inundation of salt marshes during the night followed by draining during a falling tide during the day lead to larger respiration values.  Resolving these patterns may simply require better data.  For the weighted regression approach, sustained periods of synchrony between the solar and tidal cycles presents a problem because collinearity between the two diminishes performance.  Moreover, \ac{DO} time series may reflect changes in metabolism associated with tides that are true biological effects, rather than an artifact of advection.  For example, respiration may be higher on falling tides as compared to rising tides if outwelling substrates on the falling tide support increased metabolism \citep{Sasaki09}. Although one might hope that an improved ability to resolve temporal patterns in metabolism would make it possible to resolve such changes using only a \ac{DO} time series, we suggest that this is still not possible. The open water metabolism method has generally been applied at the time scale of one day, rather than hourly. We suggest, therefore, that the weighted regression method may improve estimates of daily metabolism, but will not make it possible to resolve intraday variations in metabolism.
